---
title: "API Test"
author: "Randy Yee (pcx5@cdc.gov)"
date: "October 27, 2020"
output: html_document
---

```{r setup, include=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
```

## Get saved account parameters.

Create a json file in folder secrets of your working directory.

```{r}
secrets <- read_json("./secrets/datim.json")
user <- secrets$dhis$username
password <- secrets$dhis$password
```

## URL parts

```{r}
base_url <- "https://www.datim.org/api/29"
analytics_url <- "https://www.datim.org/api/analytics?"
analytics_dx <- "dimension=dx:"
analytics_pe <- "dimension=pe:"
analytics_co <- "dimension=co"
```


## Test API connection

```{r}
httr::set_config(httr::config(http_version=0))

res <- httr::GET(base_url,
           authenticate(user,password),
           timeout(60))

res$status_code == 200 # 200 TRUE, request fulfilled

# Status Codes
#Informational responses (100-199),
#Successful responses (200-299),
#Redirects (300-399),
#Client errors (400-499),
#and Server errors (500-599)

api_info <- jsonlite::fromJSON(httr::content(res, as="text"))
api_info <- as.data.frame(api_info$resources)
api_info
```

## Basic data call test
dimension=dx: (data)
dimension=pe: (period)

Example of API url:
https://www.datim.org/api/analytics?dimension=dx:owIr2CJUbwq&dimension=pe:QUARTERS_THIS_YEAR&filter=ou:ybg3MO3hcf4

https://www.datim.org/api/analytics?dimension=dx:Hyvw9VnZ2ch;ebCEt4u78PX&dimension=pe:LAST_12_MONTHS;LAST_FINANCIAL_YEAR&filter=ou:XOivy2uDpMF&displayProperty=SHORTNAME&skipMeta=true&includeNumDen=true

Note: filter for one OU using the UID

dimensions: dx > pe > co > ou

```{r}
test <- GET("https://www.datim.org/api/29/analytics?dimension=dx:c03urRVExYe;qkjYvdfOakY;ebCEt4u78PX;Hyvw9VnZ2ch&dimension=pe:LAST_4_QUARTERS&filter=ou:ybg3MO3hcf4&displayProperty=SHORTNAME&skipMeta=true&includeNumDen=true", # dimensions: dx - dataelement codes, pe - periods; sep by ";" within dimensions and "&" between dimensions
            authenticate(user,password),
            timeout(60))

test$status_code == 200

data <- fromJSON(rawToChar(test$content))
names(data)
data$headers
data$metaData
data$rows

df <- as.data.frame(data$rows) # Values are in rows
colnames(df) <- data$headers$column # Column names contained here
df$Value <- as.numeric(as.character(df$Value)) # Change value to numeric

df
```

=================================================================
## Build master metadataset for API calls and data formatting
All metadata codes are public.

Example code dataset URL:
https://www.datim.org/api/sqlViews/DotdxKrNZxG/data.json?var=dataSets:jKdHXpBfWop

Note: this JSON output is just one of 95 pages (pagecount)!

```{r}
dataset_codes <- GET("https://www.datim.org/api/dataSets")
dataset_codes <- jsonlite::fromJSON(rawToChar(dataset_codes$content))

master_meta <- data.frame()
for (code in dataset_codes$dataSets$id){
  url_dataset <- paste0("https://www.datim.org/api/sqlViews/DotdxKrNZxG/data.json?var=dataSets:", code)
  dataset_container <- GET(url_dataset)
  dataset_container <- jsonlite::fromJSON(rawToChar(dataset_container$content))
  
  is.null(dataset_container$pager$pageCount) # Check if there are additional pages
  # TODO: finish this case
  
  #Create a DF with the metadata
  if(length(dataset_container$listGrid$rows)!=0){
    dataset_container2 <- as.data.frame(dataset_container$listGrid$rows)
    colnames(dataset_container2) <- dataset_container$listGrid$headers$column
    
    #Append metadatasets
    master_meta <- bind_rows(master_meta, dataset_container2)
  }
}
master_meta
```

## How to specify results, targets, and narratives using data element groups

# 1) Start with dataset id

```{r}
result_targets_codes <- jsonlite::fromJSON(rawToChar(GET("https://www.datim.org/api/dataElementGroups")$content))
result_targets_codes <- as.data.frame(result_targets_codes$dataElementGroups)
result_targets_codes <- result_targets_codes %>% 
  filter(str_detect(displayName, "Results|Targets|Narratives")) %>%
  filter(str_detect(displayName, "2018|2019|2020"))
result_targets_codes
```

# 2) Use dataset id to get dataelements

```{r}
# Select dataset type id from above
dx_string <- result_targets_codes[result_targets_codes$displayName=="2020 MER Results",1]

dataelementgroup_url <- paste0("https://www.datim.org/api/dataElementGroups/",dx_string)
dataelementgroup <- GET(dataelementgroup_url,
    authenticate(user,password),
    timeout(60))
dataelementgroup <- jsonlite::fromJSON(rawToChar(dataelementgroup$content))
dataelementgroup_df <- as.data.frame(dataelementgroup$dataElements)
is.null(dataelementgroup$pager$pageCount) # Check if there are additional pages
dataelementgroup_df

# Get dataelements to map names to groups
dataelement_url <- "https://www.datim.org/api/dataElements/"
dataelement <- GET(dataelement_url,
    authenticate(user,password),
    timeout(60))
dataelement <- jsonlite::fromJSON(rawToChar(dataelement$content))
dataelement_df <- as.data.frame(dataelement$dataElements)
is.null(dataelement$pager$pageCount) # Check if there are additional pages

# Use a loop to call other pages
for (i in 2:dataelement$pager$pageCount){
  new_url <- paste0(str_sub(dataelement_url, end=-1),"?page=",i)
  dataelement2 <- GET(new_url,
    authenticate(user,password),
    timeout(60))
  dataelement2 <- jsonlite::fromJSON(rawToChar(dataelement2$content))
  dataelement_df2 <- as.data.frame(dataelement2$dataElements)
  dataelement_df <- bind_rows(dataelement_df,dataelement_df2)
}
dataelement_df

map_dataelement <- left_join(dataelementgroup_df,dataelement_df, by = "id")
map_dataelement
```

## Reference geographic codes from OU to site

```{r}
# Get orgunits to map names to groups
orgunit_url <- "https://www.datim.org/api/organisationUnits/"
orgunit <- GET(orgunit_url,
    authenticate(user,password),
    timeout(60))
orgunit <- jsonlite::fromJSON(rawToChar(orgunit$content))
orgunit_df <- as.data.frame(orgunit$organisationUnits)
is.null(orgunit$pager$pageCount) # Check if there are additional pages

# Use a loop to call other pages
for (i in 2:orgunit$pager$pageCount){
  new_url <- paste0(str_sub(orgunit_url, end=-1),"?page=",i)
  orgunit2 <- GET(new_url,
    authenticate(user,password),
    timeout(60))
  orgunit2 <- jsonlite::fromJSON(rawToChar(orgunit2$content))
  orgunit_df2 <- as.data.frame(orgunit2$organisationUnits)
  orgunit_df <- bind_rows(orgunit_df,orgunit_df2)
}
head(orgunit_df)
```

## Generic function for returning codesets

```{r}
getCodeDataset <- function(code){
  
  getcode_url <- paste0(base_url,"/",code,"?paging=FALSE")
  coded <- GET(getcode_url,
               authenticate(user,password),
               timeout(60))
  coded <- jsonlite::fromJSON(rawToChar(coded$content))
  coded_df <- as.data.frame(coded[[code]])
  names(coded_df)[names(coded_df) == 'displayName'] <- eval(code)
  return(coded_df)
  
  # if(is.null(coded$pager$pageCount)){ # Check if there are additional pages
  #   return(coded_df)
  # } else {
  #   # Use a loop to call other pages
  #   for (i in 2:coded$pager$pageCount){
  #     new_url <- paste0(str_sub(getcode_url, end=-1),"?page=",i)
  #     coded2 <- GET(new_url,
  #                   authenticate(user,password),
  #                   timeout(60))
  #     coded2 <- jsonlite::fromJSON(rawToChar(coded2$content))
  #     coded_df2 <- as.data.frame(coded2[[code]])
  #     coded_df <- bind_rows(coded_df,coded_df2)
  #   }
  #   return(coded_df)
  # }
}

getCodeDataset("dataSets")
getCodeDataset("organisationUnitGroups")
getCodeDataset("organisationUnitGroupSets")
getCodeDataset("organisationUnitLevels")
```

=====================================================================================
## Start building the URL
## BaseURL + dx dimension + pe dimension + co (category option combos) + ou + etc

Example for returning TX_CURR results for Moz and facilities for the past 4 quarters.

dimension dx: all TX_CURR dataelements
dimension pe: last 4 quarters
dimension co: include category option combos
dimension ou: for Moz and facilities
Note: (replacing dimension) filter=ou will not give you a by ou output

```{r}
pivot_url <- "https://www.datim.org/api/29/analytics?dimension=dx:c03urRVExYe;qkjYvdfOakY;ebCEt4u78PX;Hyvw9VnZ2ch&dimension=pe:LAST_4_QUARTERS&dimension=co&dimension=ou:h11OyvlPxpJ;OU_GROUP-POHZmzofoVx&displayProperty=NAME"

# See getCodeDataset("organisationUnitGroups")
# OU_GROUP-POHZmzofoVx is facility
# OU_GROUP-PvuaP6YALSA is 
# OU_GROUP-nwQbMeALRjL is

test <- GET(pivot_url,
            authenticate(user,password),
            timeout(600))

test$status_code == 200

data <- fromJSON(rawToChar(test$content))
names(data)
data$headers
data$metaData
data$rows
data$metaData$dimensions
data$metaData$items

decoder <- rownames_to_column(do.call(bind_rows, lapply(data$metaData$items, data.frame)), var = "id")

df <- as.data.frame(data$rows) # Values are in rows
colnames(df) <- data$headers$column # Column names contained here
df$Value <- as.numeric(as.character(df$Value)) # Change value to numeric

decoded_df <- df %>% 
  left_join(getCodeDataset("dataElements"), 
            by = c("Data" = "id")) %>%
  left_join(getCodeDataset("categoryOptionCombos"), 
            by = c("Category option combo" = "id")) %>%
  left_join(getCodeDataset("organisationUnits"), 
            by = c("Organisation unit" = "id"))
  
decoded_df
```

